---
title: "Class 5 In-Class Assignment"
author: "HBEH 762"
date: "September 28, 2018"
output: word_document
---

#Random intercepts model
We first fit a random intercepts model, testing the Level-1 effects of *sex* and *group-mean centered importance of relative performance* on *intrinsic motivation*

Here are the Level 1 and Level 2 equations:
##Level 1:
$intrinsic_{ij} = \beta_{0j} + \beta_{1j}sex_{ij} + \beta_{2j}relperf_{ij} + r_{ij}$
$r_{ij}\sim N\left(0,\sigma^2\right)$
##Level 2:
$\beta_{0j} = \gamma_{00} + u_{0j}$
$\beta_{1j} = \gamma_{10}$
$\beta_{2j} = \gamma_{20}$
$u_{0j}\sim N\left(0, \tau_{00}\right)$

```{r, engine="sas", engine.path="C:/Program Files/SASHome/SASFoundation/9.3/sas.exe",echo=FALSE}

libname motiv 'C:\Users\vcole\Dropbox\teaching';

data motiv;
set motiv.motivation;
run;

proc sort data=motiv;
by classid;
run;

proc means data=motiv noprint;
by classid;
var relperf GoalStrct;
output out=classmean  mean=relperf_class goalstrct;
run;

proc print data=classmean;
run;
/*calculating the grand mean*/
proc means data=classmean;
var relperf_class GoalStrct ;
run;
data classmean;
set classmean;
drop _TYPE_ _FREQ_; /*extra variables created in last step*/
relperf_class_grdmc=relperf_class-3.1690349; /*subtracting the grand mean*/
goalstrct_grdmc=goalstrct-2.8851064; /*subtracting the grand mean*/
run;

/*merge class means with original data*/
data motiv;
merge motiv classmean;
by classid;
run;

/*group-mean-center relperf*/
data motiv;
set motiv;
relperf_grpmn=relperf-relperf_class; /*subtract the un-centered class mean*/
run;

proc mixed data=motiv method=ml covtest;
class classid;
model intrinsic=sex relperf_grpmn relperf_class_grdmc goalstrct_grdmc/s;
random intercept/subject=classid ;
run;
```

#Random slopes model
```{r, engine="sas", engine.path="C:/Program Files/SASHome/SASFoundation/9.3/sas.exe",echo=FALSE}

libname motiv 'C:\Users\vcole\Dropbox\teaching';

data motiv;
set motiv.motivation;
run;

proc sort data=motiv;
by classid;
run;

proc means data=motiv noprint;
by classid;
var relperf GoalStrct;
output out=classmean  mean=relperf_class goalstrct;
run;

proc print data=classmean;
run;
/*calculating the grand mean*/
proc means data=classmean;
var relperf_class GoalStrct ;
run;
data classmean;
set classmean;
drop _TYPE_ _FREQ_; /*extra variables created in last step*/
relperf_class_grdmc=relperf_class-3.1690349; /*subtracting the grand mean*/
goalstrct_grdmc=goalstrct-2.8851064; /*subtracting the grand mean*/
run;

/*merge class means with original data*/
data motiv;
merge motiv classmean;
by classid;
run;

/*group-mean-center relperf*/
data motiv;
set motiv;
relperf_grpmn=relperf-relperf_class; /*subtract the un-centered class mean*/
run;

proc mixed data=motiv method=ml covtest;
class classid;
model intrinsic=sex relperf_grpmn relperf_class_grdmc goalstrct_grdmc/s;
random intercept relperf_grpmn/subject=classid type=un s;
run;
```


#Slopes-as-outcomes model
```{r, engine="sas", engine.path="C:/Program Files/SASHome/SASFoundation/9.3/sas.exe",echo=FALSE}

libname motiv 'C:\Users\vcole\Dropbox\teaching';

data motiv;
set motiv.motivation;
run;

proc sort data=motiv;
by classid;
run;

proc means data=motiv noprint;
by classid;
var relperf GoalStrct;
output out=classmean  mean=relperf_class goalstrct;
run;

proc print data=classmean;
run;
/*calculating the grand mean*/
proc means data=classmean;
var relperf_class GoalStrct ;
run;
data classmean;
set classmean;
drop _TYPE_ _FREQ_; /*extra variables created in last step*/
relperf_class_grdmc=relperf_class-3.1690349; /*subtracting the grand mean*/
goalstrct_grdmc=goalstrct-2.8851064; /*subtracting the grand mean*/
run;

/*merge class means with original data*/
data motiv;
merge motiv classmean;
by classid;
run;

/*group-mean-center relperf*/
data motiv;
set motiv;
relperf_grpmn=relperf-relperf_class; /*subtract the un-centered class mean*/
run;

proc mixed data=motiv method=ml covtest;
class classid;
model intrinsic=sex relperf_grpmn relperf_class_grdmc goalstrct_grdmc goalstrct*relperf_grpmn/s;
random intercept relperf_grpmn/subject=classid type=un ;
run;
```

